---
title: "Activity"
output: html_notebook
---
# Read the data and load libraries
```{r}
source('read_data.R')
```
```{r}
d <- loans %>% dplyr::sample_n(1000)
save(d,file=file.path("data","d"))
load(file=file.path("data","d"))
```
```{r}
my_count <- function(d,group_var) {
  group_var <- enquo(group_var)  # quote the value of group_var
  d %>% group_by(!!group_var) %>% 
    summarise(count = n())
}
```

```{r}
dc <- d %>% my_count(activity) %>% arrange(-count) %>%  slice(1:5)
```




Kantar

programming with dplyr
enquo  enquote
x <-  quantityu
```{r}
my_count <- function(d,group_var) {
  group_var <- enquo(group_var)  # quote the value of group_var
  d %>% group_by(!!group_var) %>% 
    summarise(count = n())
}
```


```{r}
dc <- d %>% my_count(activity) %>% arrange(-count) %>%  slice(1:5)
dc %>% ggplot() +
  aes(x=activity,y=count,group=activity,stat='identity') +
  geom_col()
choices <- list("activity","sector")
```

```{r}
inputPanel(
  selectInput("var_x", label = "Plot X variable:",
              choices = choices, selected = 'activity')
)
``` 
```{r}
renderPlot({
  dc %>% ggplot() +
  aes(x=activity,y=count,group=activity,stat='identity') +
  geom_col()
})
```


```{r}

```


```{r}
dc %>% ggplot() +
  aes(x=activity,y=count,group=activity,stat='identity') +
  geom_col()
```

```{r}
loans %>% my_count(sector)
```
```{r}
loans %>% my_count(use)
```
```{r}
loans %>% my_count(country_code)
```
```{r}
loans %>% my_count(country)
```


```{r}
loans %>% my_count(region)
```


