---
title: "Data Science for Good: Kiva Crowdfunding"
subtitle: "Code Borrower Gender"
output: html_notebook
---

Explore different ways to code the borrower gender.
The requires you have the data table/frame loans from read_data_into_R.

# Load any necessary libraries
```{r,cache=TRUE,include=FALSE}
source('library.R')
```

# Subset the Data
In the development stage, we subset the data to a smaller data set. Subset the data set to Zimbabwe. .
Make it a tibble.
```{r}
Zimbabwe <- loans %>% filter(country == "Zimbabwe") %>% 
  as.tibble()
```

# Create summary statistics
These next code chunks will demonstrate how to calculate the summaries
Here are the summaries we will we will calculate.

. Is the first borrower listed a female?
. Count of males and females
. Sum of ranks for mean and females by id

First we will take a look a the records with male borrowers and use that to check 
the subsequent calculations.

## A filter example
This is used to cross check/verify that the next queries get the correct records.
select the fields id and borrower_genders only
filter to males only
[^e]male  male not preceded by e
```{r,echo=FALSE}
gender <- Zimbabwe %>% 
  select(id,borrower_genders) %>% 
  filter(str_detect(borrower_genders, regex("[^e]male",ignore_case = TRUE))) %>% 
  arrange(id)
gender
```
## Unnest the borrower genders
Unnest example:

* Create a tidy text tibble.
* Create one record for each word.
* Select the records with male.
* Sort by id.
```{r}
Zimbabwe %>% 
  select(id,borrower_genders) %>% 
  unnest_tokens(word,borrower_genders) %>% 
  filter(word == 'male') %>% 
  arrange(id)
```
## Is the first borrower listed a female?
```{r}
gender %>% 
  mutate(female_first = str_detect(borrower_genders, regex("^female",ignore_case = TRUE))) %>% 
  group_by(id) %>%
  mutate(female_first = max(female_first)) %>% 
  arrange(id)
```

## Count of males and females

* Code a 1 for female and 1 for male 
* Count the number of females and males
* Calculate the percent of females.
```{r}
gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  group_by(id) %>%
  summarise(female = sum(str_detect(word, regex("female",ignore_case = TRUE))),
            male   = sum(str_detect(word, regex("^male",ignore_case = TRUE)))) %>% 
  mutate(female_percent = female/(female + male)) %>% 
  ungroup()
```
# Sum of ranks for men and females by id

First, show how the order/rank can be calculated.

* Code the order of male/female in borrower_genders
* Identify the position of the occurance.
* Filter for the men so the results can be checked.
```{r}
gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  mutate(female = str_detect(word, regex("female",ignore_case = TRUE))) %>% 
  group_by(id) %>% 
  mutate(order_number = row_number()) %>%
  ungroup()  %>% 
  filter(!female) %>% 
  arrange(id,order_number)
```
This is the same thing , but add a sum of ranks.
Impute missing,when there are no female or male borrowers with a zero.
```{r}
gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  mutate(female = str_detect(word, regex("female",ignore_case = TRUE))) %>% 
  group_by(id) %>% 
  mutate(order_number = row_number()) %>%
  ungroup()  %>% 
  group_by(id,word)  %>% 
  summarise(sum_rank = sum(order_number))  %>% 
  ungroup() %>% 
  spread(word,sum_rank)  %>% 
  mutate(male_rank_sum = Hmisc::impute(male,0),female_rank_sum = Hmisc::impute(female,0)) %>% 
  select(id,male_rank_sum,female_rank_sum)
```





## Put everything together
```{r}
gender_first <- gender %>% 
  mutate(gender_first = str_detect(borrower_genders, regex("^female",ignore_case = TRUE))) %>% 
  group_by(id) %>%
  mutate(gender_first = factor(max(gender_first),levels=c(0,1),labels = c("Male","Female"))) %>% 
  arrange(id) 

count_borrowers <-  gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  group_by(id) %>%
  summarise(female_count = sum(str_detect(word, regex("female",ignore_case = TRUE))),
            male_count   = sum(str_detect(word, regex("^male",ignore_case = TRUE)))) 
sum_ranks <- gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  mutate(female = str_detect(word, regex("female",ignore_case = TRUE))) %>% 
  group_by(id) %>% 
  mutate(order_number = row_number()) %>%
  ungroup()  %>% 
  group_by(id,word)  %>% 
  summarise(sum_rank = sum(order_number))  %>% 
  ungroup() %>% 
  spread(word,sum_rank)  %>% 
  mutate(male_rank_sum = Hmisc::impute(male,0),female_rank_sum = Hmisc::impute(female,0)) %>% 
  select(id,male_rank_sum,female_rank_sum)

selected_fields <- Zimbabwe %>% 
  select(id,loan_amount,funded_amount,term_in_months)

borrower_genders <-inner_join(gender_first,count_borrowers,by=c("id")) %>% 
  inner_join(sum_ranks,by=c('id')) %>% 
  inner_join(selected_fields,by=c('id'))

```
```{r}
borrower_genders
```
```{r}
save(borrower_genders,file=file.path("data","borrower_genders"))
```

```{r}
borrower_genders %>% ggplot() +
  aes(loan_amount,term_in_months,color=gender_first) +
  geom_point()

```
```{r}
borrower_genders %>% ggplot() +
  aes(loan_amount,term_in_months,fill=gender_first,group=gender_first) +
  geom_boxplot()

```
```{r}
borrower_genders %>% 
  ggplot(aes(y=loan_amount,x=gender_first,group=gender_first)) +
  geom_boxplot()

```










