---
title: "Kiva Loans"
subtitle: "Currency"
output: html_document
---

Take a look at converting the currency.

How to used the date fields

# Read the data and load libraries
```{r}
source('read_data.R')
```

```{r}
c <- loans %>% 
  select(currency,country,loan_amount,funded_amount)  %>% 
  sample_n(1000)
```

```{r}
head(c)
```

Look at currency.
I used 2014 exchange rates as given by openexchangerates.org.
I entered entered the correct url as discribed on
https://docs.openexchangerates.org/docs/api-introduction
This I saved the result to *exchangerates_2014_01_14.json*
I used the begining of 2014 as the date. This method needs 
improvement because it would be better to look at the loan date 
and use the exchange rate from that time.  But all the values 
would have to then be converted to the same date.

```{r}
echange_rates <- read_json(file.path("data","exchangerates_2014_01_14.json")) %>% 
  enter_object("rates")   %>% 
  gather_keys("currency")  %>% 
  append_values_number("rate") %>% 
  select(-document.id)
#  spread_values(rates      = jstring("rates"))
```

Next try to join this tibble to the kiva data.
```{r}
left_join(c,echange_rates) %>% 
  mutate(loan_dollars = loan_amount * rate,
         funded_dollars = funded_amount * rate)
```






