---
title: "Data Science for Good: Kiva Crowdfunding"
subtitle: "Read the Data into R"
output: html_notebook
---

# Load any necessary libraries
```{r,cache=TRUE,include=FALSE}
source('library.R')
```


```{r}
names(loans)
```
Subset the data set to Zimbabwe.
=======
# Read in the kiva_loans data.
```{r,cache=TRUE,include=FALSE}
loans <- data.table::fread(file.path("data","kiva_loans.csv"))
# This method from readr will read the csv, but it is slower than fread from data.table
#loans <- readr::read_csv(file.path("data","kiva_loans.csv"))
```

# Create summaries
>>>>>>> b8faf4451db02985ad75169aeed1de3f210b5d96
```{r}
Zimbabwe <- loans %>% filter(country == "Zimbabwe")
```

select the fields country and borrower_genders only
```{r}
gender <- loans %>% select(country, ends_with("genders"))
gender
```


```{r}
library(stringr)
```

```{r}

#Zimbabwe %>% filter(borrower_genders %in% "female") %>% 
#  select(sector,country,loan_amount,borrower_genders)

Zimbabwe %>% mutate(gender = str_detect(borrower_genders, "female")) %>% 
  select(sector,country,loan_amount,borrower_genders,gender)
```

```{r}
Zimbabwe %>% mutate(gender = grepl("^female",borrower_genders)) %>% 
  select(gender,borrower_genders,sector,country,loan_amount) %>% 
  filter(grepl("^male,",borrower_genders))
=======
```{r,cache=TRUE}
Hmisc::describe(loans)
```

# Subset the Data
Subset the data set to Zimbabwe.
Make it a tibble
```{r}
Zimbabwe <- loans %>% filter(country == "Zimbabwe") %>% 
  as.tibble()
```

# Create summary statistics
These next code chunks will demonstrate how to calculate the summaries

* Is the first borrower listed a female?
* Count of males and females
* Sum of ranks for mean and females by id

## A filter example
This is used to cross check/verify that the next queries get the correct records.
select the fields id and borrower_genders only
filter to males only
[^e]male  male not preceded by e
```{r,echo=FALSE}
gender <- Zimbabwe %>% 
  select(id,borrower_genders) %>% 
  filter(str_detect(borrower_genders, regex("[^e]male",ignore_case = TRUE))) %>% 
  arrange(id)
gender
```
## Unnest the borrower geneders
Unnest example:

* Create a tidy text tibble.
* Create one record for each word.
* Select the records with male.
* Sort by id.
```{r}
Zimbabwe %>% 
  select(id,borrower_genders) %>% 
  unnest_tokens(word,borrower_genders) %>% 
  filter(word == 'male') %>% 
  arrange(id)
```
## Is the first borrower listed a female?
```{r}
gender %>% 
  mutate(female_first = str_detect(borrower_genders, regex("^female",ignore_case = TRUE))) %>% 
  group_by(id) %>%
  mutate(female_first = max(female_first)) %>% 
  arrange(id)
```

## Count of males and females

* Code a 1 for female and 1 for male 
* Count the number of females and males
```{r}
gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  group_by(id) %>%
  summarise(female = sum(str_detect(word, regex("female",ignore_case = TRUE))),
            male   = sum(str_detect(word, regex("^male",ignore_case = TRUE)))) 
```

# Sum of ranks for mean and females by id

* Code the order of male/female in borrower_genders
* Identify the position of the occurance.
* Filter for the men so the results can be checked.
```{r}
gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  mutate(female = str_detect(word, regex("female",ignore_case = TRUE))) %>% 
  group_by(id) %>% 
  mutate(order_number = row_number()) %>%
  ungroup()  %>% 
  filter(!female) %>% 
  arrange(id,order_number)
```
Same thing, but add a sum of ranks
Impute missing,when there are no female or male borrowers with a zero.
```{r}
gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  mutate(female = str_detect(word, regex("female",ignore_case = TRUE))) %>% 
  group_by(id) %>% 
  mutate(order_number = row_number()) %>%
  ungroup()  %>% 
  group_by(id,word)  %>% 
  summarise(sum_rank = sum(order_number))  %>% 
  ungroup() %>% 
  spread(word,sum_rank)  %>% 
  mutate(male_rank_sum = Hmisc::impute(male,0),female_rank_sum = Hmisc::impute(female,0)) %>% 
  select(id,male_rank_sum,female_rank_sum)
```




Take a look at odd values of borrower_genders
=======
## Put everything together
```{r}
female_first <- gender %>% 
  mutate(female_first = str_detect(borrower_genders, regex("^female",ignore_case = TRUE))) %>% 
  group_by(id) %>%
  mutate(female_first = max(female_first)) %>% 
  arrange(id) 

count_borrowers <-  gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  group_by(id) %>%
  summarise(female = sum(str_detect(word, regex("female",ignore_case = TRUE))),
            male   = sum(str_detect(word, regex("^male",ignore_case = TRUE)))) 
sum_ranks <- gender %>% 
  unnest_tokens(word,borrower_genders) %>%
  mutate(female = str_detect(word, regex("female",ignore_case = TRUE))) %>% 
  group_by(id) %>% 
  mutate(order_number = row_number()) %>%
  ungroup()  %>% 
  group_by(id,word)  %>% 
  summarise(sum_rank = sum(order_number))  %>% 
  ungroup() %>% 
  spread(word,sum_rank)  %>% 
  mutate(male_rank_sum = Hmisc::impute(male,0),female_rank_sum = Hmisc::impute(female,0)) %>% 
  select(id,male_rank_sum,female_rank_sum)


```






