---
title: "Data Science for Good: Kiva Crowdfunding"
subtitle: "Dealing with Dates using lubridate"
output: html_notebook
---

How to used the date fields

# Read the data and load libraries
```{r}
source('read_data.R')
```


# Identify the date fields.
```{r}
Hmisc::contents(loans)
```
The time related fields are posted_time, disbursed_time, funded_time, term_in_months, repayment_interval, and date.
Look at the dates. We see that dates are a date-time string.  This can be handled with 
lubridate.

```{r}
d <- loans %>% 
  select(posted_time, disbursed_time, funded_time, term_in_months, repayment_interval,date) %>% 
  slice(1:40)
```

Note that lubridate::as_date adds the class Date to dates.  The atomic type is converted to
double. This allows numerical calculations.
```{r}
glue("The class of date is: {class(d$date)}")
glue("The type of date is: {typeof(d$date)}")
glue("The class of as_date is: {class(as_date(d$date))}")
glue("The type of as_date: is: {typeof(lubridate::as_date(d$date))}")
```


Let's convert everything to date.  Some of the fields are date time, but that level of accurace is
not really needed.  It might be interesting to see the time of day, and am or pm is track using 
the lubridate function am.
```{r}
d <- loans %>% 
  select(sector,posted_time, disbursed_time, funded_time, term_in_months, repayment_interval,date,
         funded_amount,loan_amount,term_in_months,sector) %>% 
  mutate(funded_am      = am(as_date(funded_time))) %>% 
  mutate(posted_time    = as_date(posted_time),
         disbursed_time = as_date(disbursed_time),
         funded_time    = as_date(funded_time),
         term_in_months = as_date(term_in_months),
         date           = as_date(date))
```
Let's find a good subsetting field.
*Sector* seems like a good choice.  Count the records in each sector and order
the records by descending count buy using *-sector_count*.
```{r,}
d %>% group_by(sector) %>% 
  summarize(sector_count = n()) %>% 
  arrange(-sector_count)
```
Select the *Housing* sector and sample to make it easier to plot.
```{r}
da <- d %>% 
  filter(sector %in% c('Housing','Clothing','Education')) %>% 
  sample_n(size=1000)
```

```{r}
da %>% na.omit() %>%  
  ggplot(aes(funded_time,posted_time)) + 
  geom_point() +
  geom_smooth(method=loess, se=TRUE,level=0.999) +
  labs(x="Load Funded Time",y="Loan Posted Time", title="Kiva Loan Funded and Posted Times",
       subtitle = "Posted and Funded times are almost equal",
       caption = "Kiva: Loan Times")
```

Try some date calculations.  There are *periods* *durations* and *intervals*.
```{r}
da <- da %>% 
  mutate(disbursed_2_funded = disbursed_time - funded_time)
```


Lets try this in some plots.
Interesting that loans appear to be disbursed before they are funded.
```{r}
da %>% na.omit() %>% 
  ggplot(aes(posted_time,disbursed_2_funded,color=sector)) + 
  scale_y_continuous() +
  geom_point() +
  labs(x        = "Loan Posted Date",
       y        = "Time from Funding to Dispersion", 
       title    = "Kiva Loan Posting Date and Days from Funded To Disbursion",
       subtitle = "Posted and time to Funding",
       caption  = "Kiva: Loan Funding Days")
```
Side by Side box plots.
```{r}
da %>% na.omit() %>% 
  ggplot(aes(y=disbursed_2_funded,x=sector,color=sector)) + 
  scale_y_continuous() +
  geom_boxplot() +
  labs(x        = "Loan Posted Date",
       y        = "Time from Funding to Dispersion", 
       title    = "Kiva Loan Posting Date and Days from Funded To Disbursion",
       subtitle = "Posted and time to Funding",
       caption  = "Kiva: Loan Funding Days")
```
The violin plot is an alternative to the box plot.  I think it beats a box plot.
```{r}
da %>% na.omit() %>% 
  ggplot(aes(y=disbursed_2_funded,x=sector,color=sector)) + 
  scale_y_continuous() +
  geom_violin() +
  labs(x        = "Loan Posted Date",
       y        = "Time from Funding to Dispersion", 
       title    = "Kiva Loan Posting Date and Days from Funded To Disbursion",
       subtitle = "Posted and time to Funding",
       caption  = "Kiva: Loan Funding Days")
```


```{r}
da %>% na.omit() %>%  
  ggplot(aes(funded_time,disbursed_time)) + 
  geom_point() +
  labs(x="Load Funded Time",y="Loan Disbursed Time", title="Kiva Loan Posted and Disbursed Times",
       subtitle = "Posted and Distributed times are almost equal",
       caption = "Kiva: Loan Times")
```


