---
title: "Linear Algebra"
output: html_notebook
---
```{r}
library(Matrix)
```

What is a matrix?  
How is it represented?

```{r}
A <- matrix(c(1,2,3,4,5,6),nrow=2)
A
```
Consider the case in this table

Age |  Income | Expendidure
----|---------|-------------
23  |  10     |  8
25  |  20     | 17  


```{r}
X <- matrix(c(23,10,8,25,20,17),nrow=2,byrow = TRUE)
X
```

```{r}
dim(X)
```

What is a vector?
A vector is a n by 1 matrix

```{r}
A <- c(2,3,5)
print(A)
```
## Matrix Addition
```{r}
A <- matrix(c(1,2,3,4,5,6),nrow=2,byrow = TRUE)
B <- matrix(c(2,-1,4,6,3,1),nrow=2,byrow = TRUE)
print(A)
print(B)
print(A+B)
```

The dimensions must be compatable.  The dimensions must be the same.

## Scalar Multiplication

```{r}
A <- matrix(c(2,3,4,-1,4,3),nrow=2,byrow = TRUE)
print(A)
print(3*A)
print(A/3)
```
## Combination of Operands
```{r}
A <- matrix(c(2,-1,3,2,4,9),nrow=2)
B <- matrix(c(1,0,2,3,1,2),nrow=2)
C <- matrix(c(1,6,3,7,4,5),nrow=2)
print(A)
print(B)
print(C)
print(3*B)
print(A + 3*B)
print(A + 3*B - C)
```

## Transpose
```{r}
A<-matrix(c(1,3,5,2,4,6),nrow=3)
print(A)
print(t(A))
```
Matrix Multiplication
The number of columns of A must equal the number of rows of b
The resulting matrix will have the number of columns of A and the number of rows of B.
```{r}
A <- matrix(c(2,4,1,3,5,2),nrow=3)
print(A)
B <- matrix(c(-1,4,2,2,3,1),nrow=2)
print(B)
dim(A)
dim(B)
C<-A%*%B
print(C)
dim(C)
```
```{r}
I <- matrix(c(1,1,1,1,1,1,1,1,1),nrow=3)
print(I)
print(C)
solve(C)
print(I %*% C)
I <- Matrix::diag(3)
print(I)
C %*% I
solve(C)
Matrix::solve(C,I)
```

```{r}
2*(-1) + 3*4
2*3 + 3*1
1*3 + 2*1
```
column of first matrix
row of the second.

The final matrix has the dimension of
the number of row is the first 
the number of columns in the second.


```{r}
A <- matrix(c(2,4,1,3,5,2,4,7,5,3,5,7),nrow=4)
print(A)
dim(A)
B <- matrix(c(-1,4,2,2,3,1),nrow=3)
print(B)
dim(B)
C<-A%*%B
print(C)
dim(C)
```

When we multiply A*B
The result has the number of rows in A
and the number of columns in B 
The number of columns in A and the number of row is B must be the same.

The results can be indexed by 
the row in A
the column in B 
```{r}
# 10 row 2 column 1
A[2,]
B[,1]
4*(-1) + 2*4 + 3 *2
```

y = mx + b



```{r}

```







